# Data Types and Variable Creation in R : Best Practices

>
> In this chapter,the following topics are:
>
> 1.  Introduction to data types in R, including numeric, character, logical, and factor.
>
> 2.  Naming conventions for variables in R, including using descriptive and meaningful names that follow a consistent format.
>
> 3.  Best practices for creating variables, including avoiding reserved words and using lowercase letters with underscores between words for better readability.
>
> 4.  Using the **`class()`** function to check the data type of a variable and the **`typeof()`** function to check the underlying data type.
>
> 5.  Understanding the difference between value and reference types in R.
>
> 6.  Creating and working with vectors, including using the **`c()`** function to concatenate values and the **`seq()`** function to generate sequences.
>
> 7.  Understanding the importance of data type conversions in R and using functions like **`as.numeric()`**, **`as.character()`**, and **`as.logical()`** to convert data types as needed.
>
> 8.  Tips for efficient memory management when working with large data sets, including avoiding unnecessary variable creation and removing unused variables from memory.

## Data type in R

In this chapter, we will learn about the different data types available in R, their properties, and how to work with them.

### Numeric Data Type

Numeric data type represents real numbers in R. It includes `integers`, `floating-point numbers` and `complex numbers`.

Here is an example 3.1:

```{r}
x <- 2 # integer
y <- 2.5 # float
z <- 2+ 3i # complex number
x
y
z


```

### Character Data Type

### Logical Data Type

Logical data type represents boolean values in R, which can be either **`TRUE`** or **`FALSE`**. It is created using the **`TRUE`** and **`FALSE`** keywords or the **`T`** and **`F`** shorthand.

Here is the example 3.2:

```{r}
a <- T # You can use TRUE in place of T
b <- F # You can also use FAlSE in place of F

```

### Date and Time Data Type

Date and time data type represents dates and times in R. They are created using the **`as.Date()`** and **`as.POSIXct()`** functions.

Here is an example 3.3:

```{r}
dob <- as.Date("1996-10-10")
```

### Working with Data Types in R
